Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-07-01 semantic-nd/article.tex
Next part until 4.1 (excluding).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Novelty Detection}
\label{sec:novelty-detection}
% Introduce novelty detection
Novelty detection deals with detecting that a data sample belongs to a class
originated by a distribution other than the ones the system knows about~\cite{markou2003novelty}.
It is harder than classification as only positive samples of the class are available
rendering normal classification methods unusable.

Adding capabilities to an agent to detect novel samples allows to increase its
reliability. Novelty signal can then be interpreted by the system to proceed
with more caution as its knowledge does not correctly describe reality.

% Explain threshold approach to novelty detection
Due to the nature of the sensed data which are noisy and uncertain, novelty ought to be
treated in a probabilistic way where each sample has certain probability
of being generated by a class not known to the agent and a complementary probability $P(\overline{novel}|x)$
of being generated by a known class.
The true positive and false positive rate of a novelty detection system which classifies
the set $N$ of samples as novel is given by:

@@

It leads that by extending the set $N$ with new samples, the true positive and
false positive rate can never be decreased, leading to the problem where in order
to increase detection a system needs to increase its error. This describes the
base of the \emph{error and rejection tradeoff}~\cite{chow1970optimum}, which
states that a system aiming at increasing the true-positive probability will eventually increase its
false-positive error.

This way an optimal detector can be formulated by achieving the maximum true-positive
probability without its false-positive probability increasing beyond a given limit.
This is equivalent to a \emph{continuous knapsack problem} which allows a greedy
solution by sorting the items with a value per weight function. In the case of 
detection system that can be defined as:

@@

Therefore a novelty detection system before classifying a sample $a$ as novel should (greedily)
classify any sample $b$ with a smaller cost as that would achieve a higher true positive probability 
given a fixed false positive one.

@@

Based on this, it can be said that an optimal novelty detection system is
interested in defining an order relation on all the possible inputs equivalent
to the order defined by the function: $P(\overline{novel}|x)$.
And any optimal detector can be described by the largest $P(\overline{novel}|x)$
accepted by it. This can be seen as thresholding.

Using Bayes rule and assuming a constant $P(\overline{novel})$
a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
Such a ratio is a suitable function for implementing a novelty detector system with optimal
thresholding.

@@

Note, however that in the case of dynamic graph structures, an assumption on a constant
$P(novel)$ is quite strong. And although on this first approach it is assumed to be
constant, the authors acknowledge that structure plays an important role and should
be used as prior-information when calculating $P(novel)$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Novelty Detection}
\label{sec:novelty-detection}
% Introduce novelty detection
The aim of novelty detection is to identify data samples originating from a distribution different than one of those
the system knows about~\cite{markou2003novelty}.
It is harder than classification as only positive samples of the class are available
rendering normal classification methods unusable. Adding novelty detection capabilities allows to increase 
reliability of the system. Novelty signal can be used to inform the system that it should proceed
with caution as its knowledge does not correctly describe the environment.

% Explain threshold approach to novelty detection
Due to the nature of the sensed data which are noisy and uncertain, novelty ought to be
treated in a probabilistic fashion. In such case, each sample is associated with certain probability
of being generated by a class not known to the agent and a complementary probability $P(\overline{novel}|x)$
of being generated by a known class.
The true positive and false positive rate of a novelty detection system which classifies
the set $N$ of samples as novel is given by:

@@

It follows that by extending the set $N$ with new samples, the true positive and
false positive rate can never be decreased, leading to the problem where in order
to increase detection, the system needs to increase its error. This describes the
base of the \emph{error and rejection tradeoff}~\cite{chow1970optimum}, which
states that a system aiming at increasing the true-positive probability will 
eventually increase its false-positive error.

This way an optimal detector can be formulated by achieving the maximum true-positive
probability without its false-positive probability increasing beyond a given limit.
This is equivalent to the \emph{continuous knapsack problem} which allows for a greedy
solution by sorting the items with a value per weight function. In the case of 
a detection system, this can be defined as:

@@

Therefore, a novelty detection system before classifying a sample $a$ as novel, should (greedily)
classify any sample $b$ with a smaller cost as that would achieve a higher true positive probability 
given a fixed false positive one.

@@

Based on this, it can be said that an optimal novelty detection system is aims at 
defining an order relation on all the possible inputs equivalent
to the order defined by the function: $P(\overline{novel}|x)$.
Then, the detector can be described by the largest $P(\overline{novel}|x)$ accepted which is the principle of thresholding.

Using Bayes rule and assuming a constant $P(\overline{novel})$,
a ratio between conditional and unconditional probabilities of the input $x$ is obtained.
Such a ratio is a suitable function for implementing a novelty detector system with optimal
thresholding.

@@

Note, however that in the case of dynamic graph structures, an assumption on a constant
$P(novel)$ is quite strong. Although in our first approach, it is assumed to be
constant, the authors acknowledge that structure plays an important role and should
be used as prior information when calculating $P(novel)$.
